---
title: "R Notebook"
output: html_notebook
---

Task 1.0 Plotting
```{r}
data = read.csv('dur_word_frequency.csv')
summary(data)
```
Task 1.1.
```{r}
hist(data$LogDurationA)
hist(data$LogDurationW)
```

Task 1.2.
```{r}
library(dplyr)
cond_slow <- filter(data, Cond == 'slow') %>% 
  select(LogDurationA, LogDurationW)
cond_fast <- filter(data, Cond == 'fast') %>% 
  select(LogDurationA, LogDurationW)
boxplot(cond_slow, main="Slow condition", ylim = c(-2.5, 0.7))
boxplot(cond_fast, main="Fast condition", ylim = c(-2.5, 0.7))
```
The boxplots shows, it is true, that both durations are shorter for fast speaking condition than for slow speaking condition. Also I can assert that IQR for fast condition(FC) is bigger than one for slow condition(SC), so it means that the speed of reading in slow condition approximate better to some value and in fast condition people behave more differently in reading (accroding to this sample). In addition to this, we can see that the median in SC is lesser than in FC for reading words and segments and there are linear dependency between theme (because words are parts of bigger segments).

Task 2.1 Student’s t-test.
```{r}
t.test(cond_fast$LogDurationA, mu = mean(cond_slow$LogDurationA), )
t.test(cond_fast$LogDurationW, mu = mean(cond_slow$LogDurationW), )
```
2.1.3 Interpretation

Results of t-tests provide us information, which is needed to reject the H0 - mean of word/segment duration in FC is equal to mean in SC. It is because p-value is lesser than 0.05. So we can suppose that it is true: these durations differ not only in the samples, but also in the populations. (with 95% confidence interval)

2.2 Confidence intervals
2.2.1 Explicit formula
```{r}

CI_slow <- c(mean(cond_slow$LogDurationA)-(1.96 * sd(cond_slow$LogDurationA)/length(cond_slow$LogDurationA)), mean(cond_slow$LogDurationA)+(1.96 * sd(cond_slow$LogDurationA)/length(cond_slow$LogDurationA)))
CI_slow

CI_fast <- c(mean(cond_fast$LogDurationA)-(1.96 * sd(cond_fast$LogDurationA)/length(cond_fast$LogDurationA)), mean(cond_fast$LogDurationA)+(1.96 * sd(cond_fast$LogDurationA)/length(cond_fast$LogDurationA)))
CI_fast

```

2.2.2 Function MeanCI
```{r}
library(DescTools)
MeanCI(cond_slow$LogDurationA)
MeanCI(cond_fast$LogDurationA)
```

2.2.3 Function t.test

```{r}
t.test(cond_slow$LogDurationA, mu = mean(cond_fast$LogDurationA), )[4]
t.test(cond_fast$LogDurationA, mu = mean(cond_slow$LogDurationA), )[4]
```
 Yes, these results coincide with ones, calculated with library.
 
 2.2.4 Different confidence level
```{r}
MeanCI(cond_slow$LogDurationA, conf.level = 0.99)
MeanCI(cond_fast$LogDurationA, conf.level = 0.99)
```
 As it was expected these intervals seem also to ones from 2.2.1, but there are a bit wider.  



2.3 ANOVA

2.3.1 Hypothesis
The null hypothesis consiste in that Exponent with -en, -n will be equal, because of 'n' presence and vowel shouldn't have a great impact.

2.3.2. aov()

```{r}
cond_slow <- filter(data, Cond =='slow')
summary(aov(formula = cond_slow$LogDurationA ~ cond_slow$Exponent ))
```

